---
title: "Machine Learning - Day 1"
author: "Mariachiara Fortuna"
subtitle: Exploratory analysis
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r, eval = F, echo = F}
knitr::opts_chunk$set(eval=F)
```


# Libraries

```{r}
library(dplyr)
library(ggplot2)
library(here)

```





# DATA IMPORT

```{r}

dc <-  file.path(here(), "data", "dc-wikia-data.csv") %>%
  read.csv(na.strings = "")

```


# DATA EXPLORATION

## Dataset structure

```{r}
str(dc)
```


## Summary information

```{r}
summary(dc)
```


## Simple stats

### Range

```{r}
min(dc$YEAR)

min(dc$YEAR, na.rm = T)

max(dc$YEAR, na.rm = T)

range(dc$YEAR, na.rm = T)
```


### Mean and Sd

```{r}
mean(dc$APPEARANCES, na.rm = T)

sd(dc$APPEARANCES, na.rm = T)
```


## Frequency tables

### Univariate

```{r}
table(dc$EYE)
```

### Bivariate

```{r}
ID_SEX_freq  <- table(dc$ID, dc$SEX)

ID_SEX_freq
```


```{r}
margin.table(ID_SEX_freq, margin = 1)
```



```{r}
prop.table(ID_SEX_freq)
```

```{r}
prop.table(ID_SEX_freq, margin = 2)
```


## DATA MANIPULATION

### Create a new column: Mutate

```{r}

max_year <- max(dc$YEAR, na.rm =T)

dc <-  dc %>%
  mutate(active_years = max_year - YEAR) %>%
  arrange(desc(active_years))
  
```


#### Recode factor

```{r}

dc <- dc %>% 
  mutate(
    sex = case_when(SEX == "Female Characters" ~ "F",
                    SEX == "Male Characters" ~ "M",
                    TRUE ~ "other")
    )

```



### Subsetting data (rows and columns)

```{r}

the_bad <- dc %>%
  filter(ALIGN == "Bad Characters") %>%
  select(name, ID, SEX, ALIVE, YEAR, APPEARANCES, FIRST.APPEARANCE, active_years)

```


### More complex summary stats

```{r}

the_bad %>%
  group_by(SEX) %>%
  summarize(n = n(),
            avg_appearance = mean(APPEARANCES, na.rm = T),
            avg_active_year = mean(active_years, na.rm = T))

```


### Percentage

```{r}

dc %>%
  group_by(SEX, ALIVE) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))

```




## DATA VISUALIZATION


### Barplot

```{r}
ggplot(data=dc, aes(x = SEX)) +
  geom_bar()
```


#### Barplot with grouping variable 

```{r}
ggplot(data=dc, aes(x = ALIVE, fill = SEX)) +
  geom_bar()
```


```{r}
ggplot(data=dc, aes(x = YEAR, fill = SEX)) +
  geom_bar()
```


#### Percent values

```{r}

YEAR_SEX_tab <- dc %>%
  group_by(YEAR, SEX) %>%
  summarize(n = n())

margin <- dc %>%
  group_by(YEAR) %>%
  summarize(tot = n())

perc_ys <- YEAR_SEX_tab %>%
  inner_join(margin) %>%
  mutate(perc = (n/tot)*100)


ggplot(data=perc_ys, aes(x = YEAR, y = perc, fill = SEX)) +
  geom_bar(stat = "identity")


```


###  Pie chart

```{r}

eye_tab <- dc %>%
  filter(EYE != "") %>%
  group_by(EYE) %>%
  summarize(count = n())

 ggplot(data=eye_tab, 
       aes(x=factor(1), y = count, fill = factor(EYE))) +
   geom_bar(width = 1, stat = "identity") +
   coord_polar(theta = "y")

```

### Matrix plot

```{r}
ggplot(data = dc, aes (x = EYE, y = HAIR)) +
  geom_count() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



### Histogram

```{r}
ggplot(data = dc, aes(x = YEAR)) +
  geom_histogram()
```

```{r}
ggplot(data = dc, aes(x = YEAR)) +
  geom_density()
```


### Line plot

```{r}

year_sex <- dc %>%
  group_by(YEAR, sex) %>%
  summarise(n = n())

ggplot(data = year_sex, aes(x = YEAR, y = n, color = sex)) +
  geom_line()
```



### Scatterplot 


```{r}
ggplot(data = dc %>% filter(APPEARANCES<500), 
       aes(x = active_years, y = APPEARANCES)) +
  geom_point()
  
```


```{r}
ggplot(data = dc %>% filter(APPEARANCES<500), 
       aes(x = YEAR, y = log(APPEARANCES), color = ALIVE)) +
  geom_point()
```


```{r}
ggplot(data = dc %>% filter(APPEARANCES<500 & APPEARANCES>20), 
       aes(x = YEAR, y = APPEARANCES, color = SEX)) +
  geom_point()
```


### Carpet plot

```{r}

sex_alive <- dc %>%
  group_by(SEX, ALIVE) %>%
  summarize(avg_app = mean(APPEARANCES, na.rm = T))

ggplot(data = sex_alive, aes(x = SEX, y = ALIVE)) +
  geom_tile(aes(fill = avg_app)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

